---
title: '[MODY_GENES] Replication Healthy Nevada Project'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
monofont: 'Source Code Pro'
monofontoptions: "Scale=0.7"
always_allow_html: true
params:
  alpha_threshold: 0.05
  cache: no
  dpi: 180
  gg_fontsize: 9
  output_code: no
  input_file: '/HealthyNevadaProject/HNP_carriers_8genes_EXP-085-v3_ht322_20200306.tsv'
  ALL_DATA: './Data/03_Replic_HealthyNevada/ALL_DATA.rds'
  DATA_QCed: './Data/03_Replic_HealthyNevada/DATA_QCed.rds'
  res_dir: './Data/03_Replic_HealthyNevada/MiST_results.xlsx'
output:
  bookdown::html_document2:
    theme: simplex
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
    fig_width: 6.3
    fig_height: 4.7
    number_sections: true
    self_contained: true
    mathjax: default
references:
- DOI: 10.1002/gepi.21717
  URL: 'http://dx.doi.org/10.1002/gepi.21717'
  author:
  - family: Sun
    given: Jianping
  - family: Zheng
    given: Yingye
  - family: Hsu
    given: Li
  container-title: Genetic Epidemiology
  id: sun_unified_2013
  issn: 07410395
  issue: 4
  issued:
    day: 9
    month: 3
    year: 2013
  number: 4
  page: 334--344
  publisher: Wiley
  title: A Unified Mixed-Effects Model for Rare-Variant Association in Sequencing
    Studies
  type: article-journal
  volume: 37
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(stringsAsFactors = FALSE)

script_name <- "03_Replic_HealthyNevada"
project_directory <- "./"
working_directory <- gsub("/PROJECT/", "/DATATMP/", project_directory)
output_directory <- paste(working_directory, script_name, sep = "/")

library(parallel)
library(scales) ## scale
library(readr) ## read_tsv and cols
library(writexl)
library(readxl)
library(dplyr) ## %>% 
library(purrr) # map 
library(knitr) # kable 
library(kableExtra) # styling 
library(tidyr) ## pivot
library(tidyverse)
SourceDir <- paste0(project_directory, "/utils")
source(paste0(SourceDir, "/MiST.R"))
source(paste0(SourceDir, "/Summary_clinique.R"))
source("./utils/all_my_snippets.R")

# knitr::opts_chunk$set(echo = TRUE)
### Set knitr rmarkdown chunk options
knitr::opts_chunk$set(
  results = "asis",
  size = "small",
  include = TRUE,
  echo = params$output_code,
  warning = params$output_code,
  message = params$output_code,
  tidy = FALSE,
  crop = TRUE,
  autodep = TRUE,
  fig.align = "center",
  fig.pos = "!h",
  dpi = params$dpi,
  fig.path = paste0(tempdir(), "/cache/"),
  cache = params$cache,
  cache.path = paste0(tempdir(), "/cache/")
)

### Define theme
theme_set(theme_light(base_size = params$gg_fontsize))

## read RAW etc
ALL_DATA <- read_rds(path = params$ALL_DATA)
Genes_Variants <- read_rds(path = paste0(output_directory, "/Genes_Variants.rds"))
genes_of_interest <- sort(names(Genes_Variants))
covariates_names <-  c("age_now", "is_female", "median_BMI", "PC1", "PC2", "PC3", "PC4", "PC5") 
res <- readxl::read_xlsx(path = params$res_dir) %>% 
  mutate(SubClusters = NULL)
```

# Phenotype

+ The raw file used for this analyses is ``r basename(params$input_file)`` sent by Amélie (2020/03/08).
+ We have defined the phenotype of interest `T2D_status` as 1 if `HNP_T2D_case`==1 or 0 if `HNP_T2D_control`==1.
+ We have recoded the covariate `is_female` as TRUE if `SEX`==`F` or FALSE if `SEX`==`M`.
+ We have used the following covariates as adjustement: **`r paste_and_withquote(covariates_names)`**

```{r PhenotypeQuali, results='asis'}
## Quali var

with(ALL_DATA, table(T2D_status, useNA = "always")) %>%
  broom::tidy() %>% 
  tidyr::spread(key = T2D_status, value = n) %>% 
  rename("NA" = `<NA>`) %>% 
  kable(align = "c", format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE, position = "center") %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c("T2D_status" = 3))
cat("\n\n")
with(ALL_DATA, table(is_female, useNA = "always")) %>%
  broom::tidy() %>% 
  tidyr::spread(key = is_female, value = n) %>% 
  rename("NA" = `<NA>`) %>% 
  kable(align = "c", format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE, position = "center") %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c("is_female" = 3))
cat("\n\n")
with(ALL_DATA, table(T2D_status, is_female, useNA = "always")) %>%
  broom::tidy() %>% 
  tidyr::spread(key = is_female, value = n) %>% 
  rename("NA" = `<NA>`) %>% 
  kable(align = "c", format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE, position = "center")  %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" ", "is_female" = 3))
cat("\n\n")
```

```{r PhenotypeQuanti}
## Quanti var
df.num <- ALL_DATA %>%
  dplyr::select(., covariates_names) %>%
  dplyr::select(., -is_female) %>%
  dplyr::summarise_each(
    list(
      ~ n(),
      ~ min(., na.rm = TRUE),
      ~ median(., na.rm = TRUE),
      ~ max(., na.rm = TRUE),
      ~ mean(., na.rm = TRUE),
      ~ sd(., na.rm = TRUE),
      ~ sum(is.na(.))
    )
  )  
df.num.tidy <- df.num %>% 
  gather(Variable, val) %>%
  mutate(stat = map_chr(.x = Variable, .f = ~ tail(unlist(strsplit(x = .x, "_")), 1))) %>%
  mutate(Variable = map_chr(.x = Variable, .f = ~ paste(head(unlist(strsplit(x = .x, "_")),-1), collapse = "_"))) %>%
  spread(stat, val) %>% 
  rename("NA" = "sum")

df.num.tidy %>% 
  select(Variable, n, min, median, max, mean, sd, `NA`) %>% 
  mutate_if(.predicate = is.numeric, .funs = prettyNum, digits = 4) %>% 
  kable(align = "c", format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE, position = "center") %>% 
  column_spec(1, bold = TRUE)
cat("\n\n")
```


# Material & Methods

## Ethnicity

```{r ethnicity}
cat("\n")
ggplot(
  data = ALL_DATA, 
  mapping = aes(x = PC1, y = PC2)
) +
  geom_point(shape = 21) +
  labs(title = "PCA on UK BioBank population") 
cat("\n")
```

## Genotypes

+ Genotype was set to:
  - `0`, if no pathogenic variant was detected,
  - `1`, if a pathogenic variant was detected.
  
## Quality Control

+ **[QC01]** Filter on phenotypes.  
  Any missing values in the selected phenotype and covariates make the individual excluded from the analysis.  
  Remove ``r length(read_rds(path = paste0(output_directory, "/QC1_missingValues.rds")))`` individuals.


## Statistical Analyses 
 
Rare variants have been analysed with the method *MiST* [@sun_unified_2013].  
This method has been improved in the R package [`MiSTr`](https://github.com/mcanouil/MiSTr) done by the UMR1283's biostat team.

All variants are analysed as a single cluster (at the gene scale, transcript by transcript) without subcluster based on mutation's type.  
This method provides a score $\pi$ which represents the mean effect of the cluster and a score $\tau$ which represents the heterogeneous effect of the cluster.  
The overall p-value tests the association between the set of variants and the trait of interest.  

Let the equation of the model be: $Y = \alpha X + \hat{\pi} GZ$, where

$Y$ is the phenotype matrix (n x 1) for n individuals,  
$X$ is the matrix of covariates (n x p) with p covariates,  
$Z$ is a vector of q ones for the q variants and  
$G$ is the genotype matrix (n x q) coded 1 for carriers and 0 for non-carriers.
<!-- coded 0, 1 and 2 for AA, Aa, aa, where A is a major allele and a is a minor allele. -->


# Results {.tabset .tabset-fade .tabset-pills}

Global MiST results have been written in the file `MiST_results.xlsx`. 
For details of each analysis, a file have been written for each gene (named as follow `[gene_name].xlsx`) to show all variants involved in the analysis and counts of carriers in cases and controls.


## Tables {-}

```{r Results}
res %>% 
  mutate(
    p.value.overall = cell_spec(
      formatC(p.value.overall), 
      color = ifelse(p.value.overall < 0.05 & !is.na(p.value.overall), "white", "grey"), 
      background = ifelse(p.value.overall < 0.05& !is.na(p.value.overall), "firebrick", "None"),
      bold = ifelse(p.value.overall < 0.05 & !is.na(p.value.overall), TRUE, FALSE)), 
    p.value.S.pi = formatC(p.value.S.pi),
    p.value.S.tau = formatC(p.value.S.tau),
  ) %>% 
  kable(align = "c", format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE) %>% 
  column_spec(1, bold = TRUE) 
```

## Figures {-}

```{r graphs_rare}
res_rare_signif <- res %>%  
  mutate(
    Pvalue_Overall = as.factor(
      ifelse(test = p.value.overall <= params$alpha_threshold & !is.na(p.value.overall), 
             yes = paste0("Pv<=", params$alpha_threshold), 
             no = paste0("Pv>", params$alpha_threshold, " or NA")
            )
    )
  ) %>% 
  group_by(Gene_Name) %>% 
  mutate(avg_spi = rep(mean( S.pi ), n())) %>% 
  ungroup() %>% 
  arrange(desc(avg_spi), Gene_Name, desc(S.pi)) %>%
  mutate(Gene_Name = factor(x = Gene_Name, levels = unique(Gene_Name)))
        
        
if (nrow(res_rare_signif) > 50) {
  caption.txt <- paste(
    comma(nrow(res_rare_signif)), 
    "genes were significant, only the top", params$max_signif, "is shown."
  )
  res_rare_signif <- res_rare_signif[head(order(res_rare_signif[["S.pi"]], decreasing = TRUE), params$max_signif), ]
} else {
  caption.txt <- ""
}
      
factor_norm <- with(res_rare_signif, (max(S.tau) / max(S.pi)))
S.tau_breaks <- pretty_breaks(n = 2)(res_rare_signif[["S.tau"]])
      
(ggplot(res_rare_signif) +
  geom_hline(
    data = data.frame(y = S.tau_breaks),
    mapping = aes(yintercept = y / factor_norm),
    linetype = 2, 
    size = 0.25, 
    colour = "grey87"
  ) +
  geom_bar(mapping = aes(x = Gene_Name, y = S.pi, fill = Pvalue_Overall), stat = "identity") +
  geom_bar(mapping = aes(x = Gene_Name, y = S.tau / factor_norm, fill = "S.tau"), width = 0.25, stat = "identity") +
  theme(
    legend.text = element_text(size = rel(0.9)),
    legend.key.size = unit(1, "lines"),
    legend.position = "top", 
    legend.justification = "left", 
    legend.direction = "horizontal",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.background = element_rect(colour = "grey70", fill = NA),
    axis.text.x = element_text(angle = 75, vjust = 1, hjust = 1, size = rel(0.90))
  ) +
  scale_y_continuous(
    expand = expand_scale(mult = c(0, 0.10)),
    limits = c(0, NA),
    sec.axis = sec_axis(
      trans = ~. * factor_norm,
      name = expression(Score(tau)),
      breaks = S.tau_breaks
    )
  ) +
  scale_fill_viridis_d(
    begin = 1 / 5,
    breaks = c(paste0("Pv<=", params$alpha_threshold), paste0("Pv>", params$alpha_threshold, " or NA"), "S.tau"),
    labels = list(
      bquote(Score(pi) ~ " with " ~ p[overall] <= .(params$alpha_threshold)),
      bquote(Score(pi) ~ " with " ~ p[overall] > .(params$alpha_threshold) ~ " or NA"),
      bquote(Score(tau))
    )
  ) +
  labs(
    title = paste("Rare Analyses of T2Dstatus in UKBioBank"),
    caption = caption.txt,
    y = expression(Score(pi)),
    fill = NULL
  )
) %>% 
  print()
cat("\n\n")

```

# Gene by Gene 

```{r annotation}
cat("## Clinical Data {-}\n")
ClinicalData <- read_excel(path = paste0( output_directory, "/", genes_of_interest[1], ".xlsx"), sheet = 2)  

ClinicalData %>% 
  kable(align = "c", format = "html", escape = FALSE, caption = "Clinical count same for all genes analysed.") %>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE) %>% 
  column_spec(1, bold = TRUE) 
cat("\n\n")

for (igene in genes_of_interest){
  cat(paste("## ", igene, " {- .tabset .tabset-fade .tabset-pills}\n"))
  
  MyAnnotation <- read_excel(path = paste0( output_directory, "/", igene, ".xlsx"), sheet = 1)  %>% 
    mutate(Variant = as.factor(as.character(Variant)))
  
  cat("### Annotation {-}\n")
  (ggplot(data = MyAnnotation, mapping = aes(x = Variant, y = G_1)) + 
    geom_bar(stat="identity") + 
    labs(x = "Variants", y = "Carriers", title = NULL) +
    theme(axis.text.x = element_text(angle = 65, hjust = 1, vjust = 1, size = 8))
  ) %>% print()
  cat("\n\n")

  cat("### Variants list {-}\n")
  cat(
    "Following pathogenic variants were involved in this analysis:\n", 
    paste("\n+ ", unlist(Genes_Variants[igene]), "\n"),
    sep = ""
  )
  cat("\n\n")
}

```


# R session information

```{r session-info, results = "markup"}
options("width" = 110)
sessioninfo::session_info()
```
