---
title: 'Multigene analysis Healthy Nevada Project'
author: 'Mathilde Boissel'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
monofont: "Source Code Pro"
monofontoptions: "Scale=0.7"
params:
  outputCode: FALSE
  nCores: !r parallel::detectCores()
  theme_dark: FALSE
  cache: FALSE
  dpi: 120
  ggFontSize: 9
  yname: "T2D_status"
  input_file: '/HealthyNevadaProject/HNP_carriers_8genes_EXP-085-v3_ht322_20200306.tsv'
  DATA_QCed: './Data/03_Replic_HealthyNevada/DATA_QCed.rds'
  Genes_Variants03: './Data/03_Replic_HealthyNevada/Genes_Variants.rds'
  Genes_Variants02: './Data/02_Replic_UKBioBank/Genes_Variants.rds'
  res_MiST: './Data/06_Multigene_analysis_HNP/stat_MiST.RDS'
  res_rare: './Data/06_Multigene_analysis_HNP/stat_rare.RDS'
  res_detail: './Data/06_Multigene_analysis_HNP/multigene_HNP.xlsx'
output:
  bookdown::html_document2:
    theme: simplex
    toc: true
    toc_depth: 2
    toc_float: 
      collapsed: false
    fig_width: 6.3
    fig_height: 4.7
    number_sections: true
    self_contained: true
    mathjax: default
  bookdown::pdf_document2:
    toc: true
    toc_depth: 2
    fig_width: 6.3
    fig_height: 4.7
    number_sections: true
    latex_engine: "pdflatex"
references:
- id: sun_unified_2013
  title: A Unified Mixed-Effects Model for Rare-Variant Association in Sequencing Studies
  author:
    - family: Sun
      given: Jianping
    - family: Zheng
      given: Yingye
    - family: Hsu
      given: Li
  container-title: Genetic Epidemiology
  volume: 37
  URL: 'http://dx.doi.org/10.1002/gepi.21717'
  DOI: 10.1002/gepi.21717
  issue: 4
  number: 4
  issn: 07410395
  publisher: Wiley
  page: 334--344
  type: article-journal
  issued:
    year: 2013
    month: 3
    day: 9
---

```{r setup, include = FALSE}
options(stringsAsFactors = FALSE)

script_name <- "06_Multigene_analysis_HNP"
project_directory <- "./"
working_directory <- gsub("/PROJECT/", "/DATATMP/", project_directory)
output_directory <- paste(working_directory, script_name, sep = "/")

### Load packages
invisible(
  lapply(
    c(
      "utils", "grDevices", "graphics", "stats", "methods",
      "parallel", "scales", "readxl", "mongolite", "tidyverse",
      "knitr", "Hmisc", "kableExtra"
    ),
    library,
    character.only = TRUE
  )
)
SourceDir <- paste0(project_directory, "/utils")
source(paste0(SourceDir, "/pretty_kable.R"))
source(paste0(SourceDir, "/colFmt.R"))
source("./utils/all_my_snippets.R")

### Set knitr rmarkdown chunk options
opts_chunk$set(
  results = "asis",
  size = "small",
  include = TRUE,
  echo = params$outputCode,
  warning = params$outputCode,
  message = params$outputCode,
  tidy = FALSE,
  crop = TRUE,
  autodep = TRUE,
  fig.align = "center",
  fig.pos = "!h",
  dpi = params$dpi,
  fig.path = paste0(tempdir(), "/cache/"),
  cache = params$cache,
  cache.path = paste0(tempdir(), "/cache/")
)


### Define theme
theme_set(theme_light(base_size = params$ggFontSize))

## read RAW etc
DATA_QCed <- read_rds(path = params$DATA_QCed)
Genes_Variants03 <- readr::read_rds(path = params$Genes_Variants03)
Group6 <- names(Genes_Variants03)
Genes_Variants02 <- readr::read_rds(path = params$Genes_Variants02)
Group7 <- names(Genes_Variants02)
covariates_names <- c("age_now", "is_female", "median_BMI", "PC1", "PC2", "PC3", "PC4", "PC5") 

```

# Parameters

  + "Super-Clusters"  
Two clusters of genes have been analysed :  
    - **Group6** ("8 gènes replication seen in HNP"): `r paste(Group6,collapse = ", ")`.  
    - **Group7** ("7 gènes replication seen in UK"): `r paste(Group7,collapse = ", ")`.  


# Material & Methods

## Phenotype

+ The raw file used for this analyses is ``r basename(params$input_file)`` sent by Amélie (2020/03/08).
+ We have defined the phenotype of interest `T2D_status` as 1 if `HNP_T2D_case`==1 or 0 if `HNP_T2D_control`==1.
+ We have used the following covariates as adjustement: **`r paste_and_withquote(covariates_names)`**

## Genotypes

+ Genotype was set to:
  - `0`, if no pathogenic variant was detected,
  - `1`, if a pathogenic variant was detected.
  
## Ethnicity

```{r ethnicity}
cat("\n")
ggplot(
  data = DATA_QCed, 
  mapping = aes(x = PC1, y = PC2)
) +
  geom_point(shape = 21) +
  labs(title = "PCA on Healthy Nevada Project population") 
cat("\n")
```

## Quality Control

+ **[QC01]** Filter on phenotypes.  
  Any missing values in the selected phenotype and covariates make the individual excluded from the analysis.  
  Remove ``r length(read_rds(path = paste0(gsub(script_name, "03_Replic_HealthyNevada", output_directory), "/QC1_missingValues.rds")))`` individuals.

## Statistical Analyses 
 
Rare variants have been analysed with the method *MiST* [@sun_unified_2013].  
**All variants are analysed as one cluster (based on groups mentioned above) with subclusters based on genes**.  
This method provides a score $\pi$ which represents the mean effect of the cluster and a score $\tau$ which represents the heterogeneous effect of the cluster.  
The overall p-value tests the association between the set of variants and the trait of interest.  


Let the equation of the model be: $Y = \alpha X + \hat{\pi} GZ$, where  
$Y$ is the phenotype matrix (n x 1) for n individuals,  
$X$ is the matrix of covariates (n x p) with p covariates,  
$Z$ is the matrix (q x c) coded 0 or 1 if the q-th variant belong or not to the c-th gene and  
$G$ is the genotype matrix (n x q) coded 1 for carriers and 0 for non-carriers.

The main statistical analysis were performed with the following covariates:

  + A case-control study for type 2 diabetes risk (`r params$yname`) adjusted for `r paste_and_withquote(covariates_names)`. 


# Results 

## MIST  

```{r res_MiST}
res_MiST <- map_df(c("Group6", "Group7"), function(igroup){
  read_rds(path = gsub(".RDS", paste0("_", igroup, ".RDS"), params$res_MiST)) %>% 
    mutate(Group = igroup)
})

res_MiST$covariates <- gsub(";", ";\n", res_MiST$covariates)

res_MiST %>% 
  select(Group, trait, covariates, sample_size, everything()) %>% 
  mutate_at(.vars = c("S.pi", "p.value.S.pi", "S.tau", "p.value.S.tau"), .funs = formatC) %>%  
  mutate(P.values_col = map(.x = p.value.overall,.f = function(.x) ifelse(as.numeric(.x)<0.05, TRUE, FALSE))) %>% 
  mutate(P.values_col = ifelse(is.na(P.values_col), FALSE, P.values_col)) %>% 
  mutate(
    p.value.overall = map2(.x = p.value.overall, .y = P.values_col, .f = function(.x, .y) {
      cell_spec(
      formatC(.x), 
      color = ifelse(.y, "white", "grey"), 
      background = ifelse(.y, "firebrick", "None"),
      bold = ifelse(.y, TRUE, FALSE))
    })
  ) %>% 
  select(-P.values_col) %>% 
  kable(align = "c", format = "html", escape = FALSE, row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>% 
  collapse_rows(columns = 1)
```

## Clinical data  {.tabset} 

```{r Clinicaldata}
for (igroup in c("Group6", "Group7")) {
  cat("### ", igroup, "{-} \n\n")
  clinicalDta <- read_xlsx(path =  gsub(".xlsx", paste0("_", igroup, ".xlsx"), params$res_detail), sheet = 2)
  pretty_kable(clinicalDta, row.names = FALSE)
}
```

## Odd Ratios  {.tabset} 

```{r res_rare}
for (igroup in c("Group6", "Group7")) {
  cat("### ", igroup, "{-} \n\n")
  res_rare <- read_rds(path =  gsub(".RDS", paste0("_", igroup, ".RDS"), params$res_rare))
  pretty_kable(res_rare, row.names = FALSE)
}

```

## Plots of clusters {.tabset} 

```{r plot_cluster}
for (igroup in c("Group6", "Group7")) {
  cat("### ", igroup, "{-} \n\n")
  
  annot <- read_xlsx(path =  gsub(".xlsx", paste0("_", igroup, ".xlsx"), params$res_detail), sheet = 1)
  
  (
  ggplot(data = annot, aes(y = Variant, x = GENE)) +
    ggtitle(paste0("Analysed cluster for ", igroup)) +
    ylab("Variants") +
    xlab("subClusters") +
    geom_tile(fill = "grey50", colour = "grey80") + 
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_discrete(expand = c(0, 0)) +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(), 
      panel.grid.major.y = element_blank(), 
      panel.grid.minor.y = element_blank()
    )
  ) %>% 
    print()
  
  cat("\n\n")
}

```

# References

