---
title: 'MODY_GENES Report Gene Monogenique CUSTOM T2D7'
subtitle: 'Group of genes: "CUSTOM_T2Dmono_CCDiabete" / "ACMG-CUSTOM" '
author: 'Mathilde Boissel'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
monofont: 'Source Code Pro'
monofontoptions: "Scale=0.7"
params:
  output_code: FALSE
  cache: FALSE
  dpi: 90
  gg_fontsize: 9
  raw_directory: './Data/00_MAKE_DATASET/'
  group_name: "CUSTOM_T2Dmono_CCDiabete" 
  type_report: 'Custom_report' ## CC_report, Quanti_report or Custom_report
  analysis_to_import: "CC_T2D7" 
  group_description_file: '/sequencing_data/Docs/Latest_info_gene.csv' 
  what: 'Report_light' # 'Report_light' # Report_allin
  res_dir: './Data/09_ACMG_CUSTOM_T2D7/000-RESULTS/'
  max_signif: 50
  alpha_threshold: 0.05
  phenotype_file: '/phenotypes_data/Phenotyes_latest.xlsx'
  QUALmean_treshold: 50
output:
  bookdown::html_document2:
    theme: simplex
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
    fig_width: 6.3
    fig_height: 4.7
    number_sections: true
    self_contained: true
    mathjax: default
  pdf_document:
    df_print: kable
    fig_height: 4.7
    fig_width: 6.3
    latex_engine: pdflatex
    number_sections: true
    toc: true
    toc_depth: 4
    keep_tex: true
references:
- id: sun_unified_2013
  title: A Unified Mixed-Effects Model for Rare-Variant Association in Sequencing Studies
  author:
  - family: Sun
    given: Jianping
  - family: Zheng
    given: Yingye
  - family: Hsu
    given: Li
  container-title: Genetic Epidemiology
  volume: 37
  URL: 'http://dx.doi.org/10.1002/gepi.21717'
  DOI: 10.1002/gepi.21717
  issue: 4
  number: 4
  issn: 07410395
  publisher: Wiley
  page: 334--344
  type: article-journal
  issued:
    year: 2013
    month: 3
    day: 9
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
options(stringsAsFactors = FALSE)

project_directory <- "./"

### Load packages and functions
library(parallel)
library(grid)
library(scales)
library(readxl)
library(ggrepel)
library(Hmisc)
library(kableExtra)
library(knitr)
library(DT)
library(mongolite)
library(tidyverse)

### Source MC's functions
source(paste0(project_directory, "/utils/format_pval.R"))
source(paste0(project_directory, "/utils/hijack.R"))
source(paste0(project_directory, "/utils/pretty_kable.R"))

### Set knitr rmarkdown chunk options
knitr::opts_chunk$set(
  results = "asis",
  size = "small",
  include = TRUE,
  echo = params$output_code,
  warning = params$output_code,
  message = params$output_code,
  tidy = FALSE,
  crop = TRUE,
  autodep = TRUE,
  fig.align = "center",
  fig.pos = "!h",
  dpi = params$dpi,
  fig.path = paste0(tempdir(), "/cache/"),
  cache = params$cache,
  cache.path = paste0(tempdir(), "/cache/")
)

### Define theme
theme_set(theme_light(base_size = params$gg_fontsize))
scale_colour_viridis_d <- hijack(scale_colour_viridis_d, option = "plasma", begin = 0, end = 4 / 5)
scale_colour_viridis_c <- hijack(scale_colour_viridis_c, option = "plasma", begin = 0, end = 4 / 5)
scale_fill_viridis_d <- hijack(scale_fill_viridis_d, option = "plasma", begin = 0, end = 4 / 5)
scale_fill_viridis_c <- hijack(scale_fill_viridis_c, option = "plasma", begin = 0, end = 4 / 5)
viridis_pal <- hijack(viridis_pal, option = "plasma", begin = 0, end = 4 / 5)

## Genes and Transcripts
all_groups <- read.table(file = params$group_description_file, header = TRUE, sep = ";") %>% 
  dplyr::filter(!gene_name %in% c("BOLA2", "BOLA2B", "DTX2P1-UPK3BP1-PMS2P11", "RMST")) 

list_transcripts <- all_groups %>% 
  dplyr::filter(get(params$group_name)==1) %>% 
  dplyr::select(gene_name, Transcript)


## LARGE or STRICT analysis
LARGEdefSNP <- c(
  "initiator_codon_variant", "synonymous_variant", 
  "stop_retained_variant", "stop_lost", "stop_gained", "missense_variant", 
  "incomplete_terminal_codon_variant", "coding_sequence_variant", 
  "splice_acceptor_variant", "splice_donor_variant", "splice_region_variant", 
  "regulatory_region_variant"
)

STRICTdefSNP <- c(
  "initiator_codon_variant", "stop_retained_variant", 
  "stop_lost", "stop_gained", "missense_variant", 
  "splice_donor_variant", "splice_acceptor_variant"
)

# LARGEdefINDEL <- c("all")

STRICTdefINDEL <- c("frameshift_variant", "inframe_insertion", "splice_acceptor_variant", "splice_donor_variant")


dir.create(path = params$res_dir, mode = "0777", showWarnings = FALSE)
images_dir <- paste0(params$res_dir, "Images_article/")
dir.create(path = images_dir, mode = "0777", showWarnings = FALSE)
workingdir <- "./Data/09_ACMG_CUSTOM_T2D7/"

CustomFile <- "./Docs/CUSTOM_files/custom_21072020.xlsx"
InfoCUSTOM <- read_excel(path = CustomFile, sheet = 1, col_names = TRUE, col_types = "text")

```

# Parameters

## Analysed group

Analyses have been made by group. The current group of studied is ``r params$group_name``.
The custom file used were ``r basename(CustomFile)``. 

## Genes and Transcripts Or Enhancers

The current group includes **`r length(unique(InfoCUSTOM$Gene))`** gene or **`r nrow(list_transcripts)`** potential transcripts.  
Each gene will be analysed at the transcript level, for all transcripts available.


## `STRICT` to `CUSTOM`  

Filter `STRICT` have been selected

  + When the analysis selected is `STRICT`  
    - SNPs were selected when: `r paste("  \n", paste0(paste(rep("  ", 4), collapse = ""), '+ "', STRICTdefSNP, '"\n'), sep = "")`.
    - INDELs were selected when: `r paste("  \n", paste0(paste(rep("  ", 4), collapse = ""), '+ "', STRICTdefINDEL, '"\n'), sep = "")`.
    
  + Then, from `STRICT` selection, the QC11 do the `CUSTOM` final selection based on the file ``r basename(CustomFile)``.  


## Other parameters

- Minor allele frequency threshold to define rare variant: $MAF < 0.01$
- Exclusion of `Synonymous` variant: `FALSE`
- Cluster in *MiST* analysis: `none`  
  i.e., all variants are analysed as a single cluster (at the transcript scale for all gene) 
  without subcluster based on mutation’s type.


# Material & Methods

## Phenotypes

```{r Phenotypes}
Phenotypes <- readxl::read_excel(path = params$phenotype_file, col_names = TRUE, guess_max = 9000) %>%
  as.data.frame %>% 
  column_to_rownames(var = "IID")
```

  + For ethnicity, we performed a principal component analysis (PCA), 
    using **15,020** SNPs available in `MODY_GENES` and `1,000 Genomes project`, with $MAF>0.05$ within both populations.


```{r ethnicity}
cat("\n")
ggplot(
  data = dplyr::filter(Phenotypes, !is.na(Pop_pred)), 
  mapping = aes(x = PC1, y = PC2, colour = Pop_pred, shape = Pop_pred)
) +
  geom_point() +
  stat_ellipse(size = 1.5, colour = "white") +
  stat_ellipse(size = 1) +
  scale_colour_viridis_d() +
  scale_shape_discrete(solid = FALSE) +
  labs(
    title = "Predicted ethnicity of the MODY_GENES population",
    subtitle = paste0(
      "(Ethnicity was not computed for ", 
      scales::comma(nrow(dplyr::filter(Phenotypes, is.na(Pop_pred)))), 
      " samples)"), 
    colour = "Predicted\nPopulation",
    shape = "Predicted\nPopulation"
  ) 
cat("\n")
```


## Statistical Analyses 
 
Rare variants have been analysed with the method *MiST* [@sun_unified_2013].  
All variants are analysed as a single cluster (at the gene scale, transcript by transcript) without subcluster based on mutation's type.  
This method provides a score $\pi$ which represents the mean effect of the cluster and a score $\tau$ which represents the heterogeneous effect of the cluster.  
The overall p-value tests the association between the set of variants and the trait of interest.  


Let the equation of the model be: $Y = \alpha X + \hat{\pi} GZ$, where
$Y$ is the phenotype matrix (n x 1) for n individuals,
$X$ is the matrix of covariates (n x p) with p covariates,
$Z$ is a vector of q ones for the q variants and
$G$ is the genotype matrix (n x q) coded 0, 1 and 2 for AA, Aa, aa, where A is a major allele and a is a minor allele. 


Statistical analyses have been performed with the following covariates:

  
```{r covariables}
cat(
  "\n  + `age`\n",
  "\n  + `sex`\n",
  ifelse(
    test = any(params$analysis_to_import %in% c("CC_Obesity", "CC_OBESITE", "CC_Obesity_child")),
    yes =  "\n  + `BMI` (not for obesity related trait)\n", 
    no = "\n  + `BMI`\n"
  ),
  ifelse(
    test = any(params$analysis_to_import %in% c("CC_Obesity", "CC_OBESITE")),
    yes =  "\n  + `INFO_DIAB` (only for CC_OBESITE and CC_Obesity with adults)\n", 
    no = ""
  ),
  "\n  + `PC1` to `PC5` (ethnicity)\n"
)
```

```{r analysis_to_import}
cat(
  "The following analys", ifelse(length(params$analysis_to_import)>1, "e", "i"), "s have been performed:  \n",
  paste("\n  + ", params$analysis_to_import, "\n"), 
  sep = ""
)
```


# Results

```{r GET_BACK_RDS}
cat("\n")

RES_RARE <- list.files(path = paste0(workingdir), pattern = "xxr.RDS", recursive = TRUE, full.names = TRUE) %>% 
  lapply(X = ., FUN = function(xrds) {
    res <- readRDS(xrds) 
    if (nrow(res$res_rare[[1]])==0) {
      tmp <- matrix(rep(NA, 14), nrow = 1, ncol = 14) 
      colnames(tmp) <- names(res$res_rare[[1]])
      qc11_done <- res$qc11_done[[1]]
      names(qc11_done) <- "qc11_done"
      bind_cols(res %>% select(-res_rare, -qc11_done), as.data.frame(tmp)) %>% 
        mutate(qc11_done = qc11_done)
    } else {
      res %>% 
      unnest(cols = c(res_rare, qc11_done)) 
    }
  }) 

Log <- list.files(path = paste0(workingdir), pattern = "xxLog.RDS", recursive = TRUE, full.names = TRUE) %>% 
  lapply(X = ., FUN = function(xrds) {
    readRDS(xrds)
  })

QC_db <- list.files(path = paste0(workingdir), pattern = "xxqc.RDS", recursive = TRUE, full.names = TRUE) %>% 
  lapply(X = ., FUN = function(xrds) {
    readRDS(xrds)
  })


#### get back results ####

res_rare <- RES_RARE %>% 
  bind_rows() %>% 
  mutate(analyse_LARGE_STRICT = "STRICT>ACMG_CUSTOM") %>%
  dplyr::select(Gene_Symbol, ENST, analyse_LARGE_STRICT, y_name, covariates, sample_size, qc11_done, SubClusters, everything()) %>% 
  mutate(SubClusters1 = NULL) %>% 
  dplyr::select(-error, error)

names(res_rare) <- gsub("\\.", "_", names(res_rare))

if (params$type_report == "Custom_report") {
  all_data <- res_rare %>% 
    filter(nb_rare_var>0) %>%  
    dplyr::filter(
      y_name %in% params$analysis_to_import
    ) %>% 
    dplyr::group_by(y_name, analyse_LARGE_STRICT) %>% 
    tidyr::nest() %>% 
    dplyr::mutate(
      data = purrr::map(
        .x = data, 
        .f = right_join, 
        y = res_rare %>% select(ENST, Gene_Symbol) %>% distinct(), 
        by = c("Gene_Symbol", "ENST")
      )
    ) %>% 
    tidyr::unnest(cols = c(data)) %>% 
    mutate(
      y_graph = OR - 1
    ) %>% 
    select(
      c("y_name","analyse_LARGE_STRICT", "Gene_Symbol", "ENST", "OR", "y_graph", "p_value_overall", "p_value_S_pi", "p_value_S_tau", "error")
    ) %>%
    group_by(y_name, analyse_LARGE_STRICT) %>% 
    tidyr::nest() %>% 
    mutate(
      main_y_axis = purrr::map(.x = data, .f = function(.data) {
        .data %>% 
          mutate(or_order = replace_na(OR - 1, 0)) %>% 
          arrange(or_order) %>%
          mutate(x_genes = factor(x = Gene_Symbol, levels = unique(Gene_Symbol))) %>% 
          select(-c("Gene_Symbol", "ENST", "or_order")) %>%
          mutate(signe = factor(sign(y_graph)))
      }),
      second_y_axis = purrr::map(.x = main_y_axis, .f = function(.data) {
        .data %>% 
          tidyr::gather(key = pvalues_name, value = pvalue, c("p_value_overall", "p_value_S_pi", "p_value_S_tau")) %>% 
          mutate(
            pval_trans = -log10(pvalue),
            pvalues_name = factor(pvalues_name, levels = c("p_value_overall", "p_value_S_pi", "p_value_S_tau")),
            factor_norm = (max(pval_trans, na.rm = TRUE) / (max(y_graph, na.rm = TRUE))),
            pval_trans_norm =  pval_trans / factor_norm
          )
      })
    ) %>% 
    mutate(
      gg = purrr::map2(.x = main_y_axis, .y = second_y_axis, .f = function(main_y, second_y) {
        ggplot(data = main_y) +
          geom_bar(
            aes(
              x = x_genes,
              y = y_graph,
              fill = signe
            ),
            stat = "identity", 
            show.legend = FALSE,
            alpha = 2/3
          ) +
          geom_point(
            data = second_y,
            aes(
              x = x_genes,
              y = pval_trans_norm,
              colour = pvalues_name,
              shape = pvalues_name,
              size = pvalues_name
            ),  
            inherit.aes = FALSE
          ) +
          geom_hline(
            aes(yintercept = -log10(0.05)/unique(second_y[["factor_norm"]]), linetype = "-log10(0.05)"),
            colour = "firebrick2"
          ) +
          labs(x = NULL, y = "OR") +
          scale_fill_viridis_d(
            option = "plasma", 
            begin = 1/5, 
            end = 4/5
          ) +
          scale_linetype_manual(
            name = "P-Value threshold", 
            values = 2, 
            label = expression(alpha==0.05)
          ) +
          scale_colour_manual(
            name = "Parameter's p-value",
            label = list("Overall", expression("Score"~pi), expression("Score"~tau)),
            values = c(
              "p_value_overall" = "firebrick2", 
              "p_value_S_pi" = "#B12A90", 
              "p_value_S_tau" = "#E16462"
            )
          ) +
          scale_shape_manual(
            name = "Parameter's p-value",
             label = list("Overall", expression("Score"~pi), expression("Score"~tau)),
            values = c("p_value_overall" = 23, "p_value_S_pi" = 25, "p_value_S_tau" = 24)
          ) +
          scale_size_manual(
            name = "Parameter's p-value",
            label = list("Overall", expression("Score"~pi), expression("Score"~tau)),
            values = c("p_value_overall" = 2, "p_value_S_pi" = 1.25, "p_value_S_tau" = 1.25)
          ) +
          scale_y_continuous(
            labels = function(x) {x+1},
            expand = expansion(mult = c(0.05, 0.05)), 
            sec.axis = sec_axis(
              trans = ~ . * unique(second_y[["factor_norm"]]),
              name = expression(italic(P)),
              labels = function(x) {
                parse(text = ifelse(x==0, "1", paste0("10^-", x)))
              },
              breaks = pretty_breaks(n = 3)(range(second_y[["pval_trans"]], na.rm = TRUE))
            )
          ) +
          theme(
            axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.8), face = "italic")
          )
      })
    ) 
    
    all_data <- all_data %>% 
    mutate(
      main_y_axis_bis = map(.x = data, .f = function(.data) {
        
        .data %>% 
          mutate(or_order = replace_na(OR - 1, 0)) %>% 
          arrange(or_order) %>% 
          mutate(
            Gene_Symbol = map2_chr(.x = error, .y = Gene_Symbol, .f = function(error, Gene_Symbol){
              res <- ifelse(
                test = grepl(x = error, pattern = "[warning] glm.fit: fitted probabilities numerically 0 or 1 occurred", perl = FALSE, fixed = TRUE),
                yes = paste0(Gene_Symbol, "*"),
                no = ifelse(test = error %in% "None", yes = Gene_Symbol, no = paste0(Gene_Symbol, "#"))
              )
              return(res)
            })
          ) %>% 
          mutate(OR = ifelse(test = !(is.na(error) | error%in% "None"), yes = NA, no = OR)) %>% 
          mutate(or_order = ifelse(test = !(is.na(error) | error%in% "None"), yes = NA, no = or_order)) %>% 
          mutate(y_graph = ifelse(test = !(is.na(error) | error%in% "None"), yes = NA, no = y_graph)) %>% 
          mutate(p_value_overall = ifelse(test = !(is.na(error) | error%in% "None"), yes = NA, no = p_value_overall)) %>% 
          mutate(p_value_S_pi = ifelse(test = !(is.na(error) | error%in% "None"), yes = NA, no = p_value_S_pi)) %>% 
          mutate(p_value_S_tau = ifelse(test = !(is.na(error) | error%in% "None"), yes = NA, no = p_value_S_tau)) %>% 
          mutate(x_genes = factor(x = Gene_Symbol, levels = unique(Gene_Symbol))) %>% 
          select(-c("Gene_Symbol", "ENST", "or_order")) %>%
          mutate(signe = factor(sign(y_graph)))
      }),
      second_y_axis_bis = map(.x = main_y_axis_bis, .f = function(.data) {
        .data %>% 
          gather(key = pvalues_name, value = pvalue, c("p_value_overall", "p_value_S_pi", "p_value_S_tau")) %>% 
          mutate(
            pval_trans = -log10(pvalue),
            pvalues_name = factor(pvalues_name, levels = c("p_value_overall", "p_value_S_pi", "p_value_S_tau")),
            factor_norm = (max(pval_trans, na.rm = TRUE) / (max(y_graph, na.rm = TRUE))),
            pval_trans_norm =  pval_trans / factor_norm
          )
      })
    ) %>% 
    mutate( 
      gg_withoutErrors = map2(.x = main_y_axis_bis, .y = second_y_axis_bis, .f = function(main_y, second_y) {
        
        ggplot(data = main_y) +
          geom_bar(
            aes(
              x = x_genes,
              y = y_graph,
              fill = signe
            ),
            stat = "identity", 
            show.legend = FALSE,
            alpha = 2/3
          ) +
          geom_point(
            data = second_y,
            aes(
              x = x_genes,
              y = pval_trans_norm,
              colour = pvalues_name,
              shape = pvalues_name,
              size = pvalues_name
            ),  
            inherit.aes = FALSE
          ) +
          geom_hline(
            aes(yintercept = -log10(0.05)/unique(second_y[["factor_norm"]]), linetype = "-log10(0.05)"),
            colour = "firebrick2"
          ) +
          labs(x = NULL, y = "OR") +
          scale_fill_viridis_d(
            option = "plasma", 
            begin = 1/5, 
            end = 4/5
          ) +
          scale_linetype_manual(
            name = "P-Value threshold", 
            values = 2, 
            label = expression(alpha==0.05)
          ) +
          scale_colour_manual(
            name = "Parameter's p-value",
            label = list("Overall", expression("Score"~pi), expression("Score"~tau)),
            values = c(
              "p_value_overall" = "firebrick2", 
              "p_value_S_pi" = "#B12A90", 
              "p_value_S_tau" = "#E16462"
            )
          ) +
          scale_shape_manual(
            name = "Parameter's p-value",
             label = list("Overall", expression("Score"~pi), expression("Score"~tau)),
            values = c("p_value_overall" = 23, "p_value_S_pi" = 25, "p_value_S_tau" = 24)
          ) +
          scale_size_manual(
            name = "Parameter's p-value",
            label = list("Overall", expression("Score"~pi), expression("Score"~tau)),
            values = c("p_value_overall" = 2, "p_value_S_pi" = 1.25, "p_value_S_tau" = 1.25)
          ) +
          scale_y_continuous(
            labels = function(x) {x+1},
            expand = expansion(mult = c(0.05, 0.05)),
            sec.axis = sec_axis(
              trans = ~ . * unique(second_y[["factor_norm"]]),
              name = expression(italic(P)),
              labels = function(x) {
                parse(text = ifelse(x==0, "1", paste0("10^-", x)))
              },
              breaks = pretty_breaks(n = 3)(range(second_y[["pval_trans"]], na.rm = TRUE))
            )
          ) +
          theme(
            axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.8), face = "italic")
          ) +
          labs(
            caption = paste0(
              "* : Warning 'fitted probabilities numerically 0 or 1 occurred'", "\n", 
              "# : Others warning or MiST not performed"
            )
          ) 
          
      })
    )
 
# all_data$gg_withoutErrors
}


```


## Rare Variants {.tabset .tabset-fade .tabset-pills}

### Figures {- .tabset}

```{r graphs_rare}
for (ianalysis in params$analysis_to_import) {
  res_rare_signifi <- res_rare %>% 
    filter(y_name %in% ianalysis & !(is.na(p_value_overall)) & p_value_overall <= params$alpha_threshold)
  
  analysis_type = unique(res_rare$analyse_LARGE_STRICT)
  
  {
    cat("\n\n#### ", ianalysis, "_", analysis_type, " {-}\n", sep = "")
    res_rare_signif <- res_rare_signifi %>% 
      filter(analyse_LARGE_STRICT %in% analysis_type)

    if (nrow(res_rare_signif) > 0) {
      res_rare_signif <- res_rare_signif %>% 
        mutate(
          Pvalue_Overall = as.factor(
            ifelse(test = p_value_overall < 0.01, yes = "Pv<0.01", no = paste0("Pv<=", params$alpha_threshold))
          ),
          Genes = paste0(Gene_Symbol, "_", ENST)
        ) %>% 
        group_by(Gene_Symbol) %>% 
        mutate(avg_spi = rep(mean(S_pi), n())) %>% 
        ungroup() %>% 
        arrange(desc(avg_spi), Gene_Symbol, desc(S_pi)) %>%
        mutate(Genes = factor(x = Genes, levels = unique(Genes)))
        
        
      if (nrow(res_rare_signif) > params$max_signif) {
        caption.txt <- paste(
          comma(nrow(res_rare_signif)), 
          "genes were significant, only the top", params$max_signif, "is shown."
        )
        res_rare_signif <- res_rare_signif[head(order(res_rare_signif[["S_pi"]], decreasing = TRUE), params$max_signif), ]
      } else {
        caption.txt <- ""
      }
      
      factor_norm <- with(res_rare_signif, (max(S_tau) / max(S_pi)))
      s_tau_breaks <- pretty_breaks(n = 2)(res_rare_signif[["S_tau"]])
      
      fig_rare <- (ggplot(res_rare_signif) +
        geom_hline(
          data = data.frame(y = s_tau_breaks),
          mapping = aes(yintercept = y / factor_norm),
          linetype = 2, 
          size = 0.25, 
          colour = "grey87"
        ) +
        geom_bar(mapping = aes(x = Genes, y = S_pi, fill = Pvalue_Overall), stat = "identity") +
        geom_bar(mapping = aes(x = Genes, y = S_tau / factor_norm, fill = "S_tau"), width = 0.25, stat = "identity") +
        theme(
          legend.text = element_text(size = rel(0.9)),
          legend.key.size = unit(1, "lines"),
          legend.position = "top", 
          legend.justification = "left", 
          legend.direction = "horizontal",
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.background = element_rect(colour = "grey70", fill = NA),
          axis.text.x = element_text(angle = 75, vjust = 1, hjust = 1, size = rel(0.90))
        ) +
        scale_y_continuous(
          expand = expansion(mult = c(0, 0.10)),
          limits = c(0, NA),
          sec.axis = sec_axis(
            trans = ~. * factor_norm,
            name = expression(Score(tau)),
            breaks = s_tau_breaks
          )
        ) +
        scale_fill_viridis_d(
          begin = 1 / 5,
          breaks = c("Pv<0.01", paste0("Pv<=", params$alpha_threshold), "S_tau"),
          labels = list(
            bquote(Score(pi) ~ " with " ~ p[overall] < 0.01),
            bquote(Score(pi) ~ " with " ~ p[overall] <= .(params$alpha_threshold)),
            bquote(Score(tau))
          )
        ) +
        labs(
          title = paste("Rare Analyses of", ianalysis),
          caption = caption.txt,
          y = expression(Score(pi)),
          fill = NULL
        )
      ) 
      fig_rare %>% 
        print()
      cat("\n")
      
      
    } else {
      cat("NOTE: No significant `Pvalue_Overall`!\n")
      cat("\n")
    }
  
    if(params$type_report == "Custom_report"){
      if(all_data %>% filter(y_name==ianalysis & analyse_LARGE_STRICT==analysis_type) %>% nrow == 0){
        
      } else {
        ggwithout <- all_data %>% 
          filter(y_name==ianalysis & analyse_LARGE_STRICT==analysis_type) %>% 
          .[["gg_withoutErrors"]] %>% 
          .[[1]] 
        
        ggwithout %>% 
          print()
          cat("\n")
   
        ggsave(
          plot =  ggwithout + 
            theme_light(base_size = 12) + 
            theme(
              legend.position = "none", 
              axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.8), face = "italic")
            ), 
          filename = paste0(
            images_dir, 
            "OR_",
            params$group_name, "_",
            ianalysis,"_",
            analysis_type, "_warning.png"
          ),
          device = "png",
          width = 6.3, 
          height = 4.7, 
          units = "in",
          dpi = 300
        ) 
    
      }
    }
  }
}

```

### Tables {-}

```{r res_rare}
if (params$what %in% "Report_allin") {
  message("Report_allin")
  DT::datatable(
    data = res_rare, 
    rownames = FALSE,
    extensions = 'Buttons', 
    options = list(
      dom = 'Bfrtip',
      buttons = c('csv', 'excel'),
      scrollX = TRUE
    )
  )
} 
if (params$what %in% "Report_light") {
  message("Report_light")
  cat(
    "`Report_light` selected: See results in excel files : `",
    paste0(
      "000-RESULTS/", 
      params$group_name, "_", 
      params$analysis_to_import,
      "_rare.xlsx"
    ), "`.\n\n"
  )
  list(res_rare) %>%
  `names<-`(c("RareResults")) %>%
  writexl::write_xlsx(
    path = paste0(
      params$res_dir, 
      params$group_name, "_", 
      params$analysis_to_import,
      "_rare.xlsx"
    ),
    col_names = TRUE
  )
}
```


## Genes/Transcripts Informations

```{r gene_by_gene, results = "asis"}
for (igene in unique(list_transcripts[["gene_name"]])) {
  list_transcripts_igene <- list_transcripts %>%
    filter(gene_name == igene) %>%
    .$Transcript

  cat("\n")
  cat("\n###", igene, "\n")
  cat("\n")

  for (itranscript in list_transcripts_igene) {
    cat("\n####", itranscript, paste0("(", igene, ")"), "{.tabset .tabset-fade .tabset-pills}\n\n")
    
    file_raw = paste0(params$raw_directory, igene, "/", igene, "_", itranscript, "_raw.xlsx")
    
    if (file.exists(file_raw)) {
      annotation_raw <- readxl::read_excel(path = file_raw, sheet = 1)
  
      ## graph commun pour ce gene, ce transcript ...
      cat("\n")
      cat("\n##### Figures {- .tabset}\n")
      cat("\n")
      
      cat("\n")
      cat("\n###### Quality and Depth {-}\n")
      cat("\n")
  
      annotation_raw <- annotation_raw %>%
        dplyr::filter(!is.na(QUALmean)) %>%
        dplyr::filter(!is.na(DPmean)) %>%
        dplyr::filter(!is.na(Type_Variant))
  
      if (nrow(annotation_raw) > 0) {
        gg_annotation <- annotation_raw %>%
          dplyr::mutate(x = as.numeric(gsub(".*_", "", chr_pos)) / 1000) %>%
          tidyr::gather(key = "quality_measure", value = "y", QUALmean, DPmean) %>%
          dplyr::select(quality_measure, x, y, Type_Variant) %>%
          dplyr::mutate(quality_measure = factor(x = quality_measure, levels = c("QUALmean", "DPmean")))
  
        p_quality <- ggplot2::ggplot(
          data = gg_annotation,
          mapping = ggplot2::aes(x = x, y = y)
        ) +
          ggplot2::geom_point(shape = 21, colour = viridis_pal(begin = 1 / 5, end = 1 / 5)(1)) +
          ggplot2::geom_hline(
            data = data.frame(yintercept = params$QUALmean_treshold, quality_measure = "QUALmean"),
            mapping = ggplot2::aes(yintercept = yintercept),
            colour = "red"
          ) +
          ggplot2::facet_grid(cols = dplyr::vars(Type_Variant), rows = dplyr::vars(quality_measure), scales = "free_y") +
          ggplot2::labs(
            x = "Position on the gene (Kb)",
            y = NULL, # "Quality/Depth Measure",
            title = paste0(
              igene, " (Chromosome ", unique(gsub("_.*", "", annotation_raw[["chr_pos"]])), ")"
            ),
            subtitle = paste0(
              scales::percent(
                x = replace_na(
                  data = sum(
                    annotation_raw[["QUALmean"]] < params$QUALmean_treshold &
                      annotation_raw[["Type_Variant"]] %in% "SNP"
                  ) / sum(annotation_raw[["Type_Variant"]] %in% "SNP"),
                  replace = 0
                )
              ),
              " of SNPs and ",
              scales::percent(
                x = replace_na(
                  data = sum(
                    annotation_raw[["QUALmean"]] < params$QUALmean_treshold &
                      annotation_raw[["Type_Variant"]] %in% "INDEL"
                  ) / sum(annotation_raw[["Type_Variant"]] %in% "INDEL"),
                  replace = 0
                )
              ),
              " of INDELs excluded because of bad quality"
            ),
            caption = paste("The red line is the QUAL threshold =", params$QUALmean_treshold)
          ) +
          ggplot2::scale_x_continuous(
            labels = scales::comma, 
            expand = expansion(mult = c(0.1, 0))
          ) +
          ggplot2::scale_y_continuous(limits = c(0, NA))
  
        print(p_quality)
        cat("\n")
      } else {
        cat("NOTE: Data not available to draw the figure!\n")
        cat("\n")
      }
        
      cat("\n")
      cat("\n###### Consequences {-}\n")
      cat("\n")
  
      if (nrow(annotation_raw) > 0 && ! is.null(annotation_raw[["Consequence"]])) {
        (
          annotation_raw[["Consequence"]] %>%
            table() %>%
            as.data.frame() %>%
            `names<-`(c("Consequence", "Count")) %>%
            dplyr::mutate(Percentage = Count / sum(Count)) %>%
            ggplot2::ggplot(mapping = ggplot2::aes(x = Consequence, y = Percentage)) +
            ggplot2::geom_bar(stat = "identity", fill = viridis_pal(begin = 1 / 5, end = 1 / 5)(1)) +
            ggplot2::geom_text(mapping = ggplot2::aes(label = Count), vjust = -0.5, colour = "black") +
            ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 60, vjust = 1, hjust = 1)) +
            ggplot2::scale_y_continuous(
              labels = scales::percent, 
              expand = expansion(mult = c(0, 0.15))
            ) +
            ggplot2::labs(
              x = NULL,
              y = "Percentage",
              title = igene,
              subtitle = "Count and percentage of consequence types available"
            )
        ) %>%
          print()
        cat("\n")
      } else {
        cat("NOTE: Data not available to draw the figure!\n")
        cat("\n")
      }
        
      ## RES specifique a chaque analyses:
      cat("\n")
      cat("\n##### QC and Clinical Data {- .tabset}\n")
      cat("\n")
      for (ianalysis in params$analysis_to_import) {
        analysis_type = unique(res_rare$analyse_LARGE_STRICT)
        {
          cat("\n")
          cat("\n###### ", ianalysis, "_", analysis_type, " {-}\n", sep = "")
          cat("\n")
          
          ## Table or Image (Flowchart?) of the QC at each step
          QC0to11 <- QC_db[grep(igene,QC_db)][[1]] %>% 
            select(starts_with("QC"))
     
          if (nrow(QC0to11) > 0) {
            lapply(QC0to11, FUN = function(x) {
               x[[1]]$Comment
            }) %>%
              do.call("rbind",.) %>%
              `colnames<-`("Comment") %>%
              as.data.frame() %>% 
              tibble::rownames_to_column(var = "Step") %>%
              dplyr::mutate(
                Step = Step %>%
                  gsub("QC", "", .) %>%
                  as.numeric() %>%
                  sprintf("QC%02d", .),
                Comment = Comment %>%
                  gsub("After[^,]*, ", "", .) %>%
                  gsub("On[^,]*, ", "", .) %>%
                  gsub("there is no", "0", .)
              ) %>%
              pretty_kable(data = ., row.names = FALSE, align = c("c", "l"), full_width = FALSE) 
            cat("\n")
          } else {
            cat("NOTE: Data not available!\n")
            cat("\n")
          }
          
          if ("clinicalData" %in% names(Log[grep(igene,Log)][[1]])) {
            clinDta <- Log[grep(igene,Log)][[1]] %>% 
              select(clinicalData) %>% 
              unlist(., recursive = FALSE)
              
            if(length(clinDta)==2) {
              ## first elecment : ClinicalData df to show
              #  sec elecment : Mean sd.. of trait per number of carried mutation
              trash <- lapply(X = clinDta, FUN = function(.x) {
                .x %>% ## sec elecment : ClinicalData df to show
                as.data.frame() %>%
                (function(dt) {
                  if (nrow(dt) == 0) {
                    cat("NOTE: No clinical data available!\n")
                    cat("\n")
                  } else {
                    pretty_kable(data = dt, row.names = FALSE, align = "c", full_width = FALSE) 
                    cat("\n")
                  }
                }) %>%
                invisible()
              })
              
              ## ggplot saved to includ 
              ggtoinclud <- paste0(
                gsub("000-RESULTS/", "", params$res_dir),
                igene, "/", 
                igene, "_", itranscript, "_", analysis_type, "_", ianalysis, "_boxplot.png"
              )
              
              if (!file.exists(ggtoinclud)) {
                cat("NOTE: No graph about trait ~ Number of carried mutations!\n")
                cat("\n")
              } else {
                cat("![](",ggtoinclud,")")
                cat("\n")
              }
              
              
            } else {
              clinDta$clinicalData %>% ## length == 1 (only ClinicalData df to show)
              as.data.frame() %>%
              (function(dt) {
                if (nrow(dt) == 0) {
                  cat("NOTE: No clinical data available!\n")
                  cat("\n")
                } else {
                  pretty_kable(data = dt, row.names = FALSE, align = "c", full_width = FALSE) 
                  cat("\n")
                }
              }) %>%
              invisible()
            }
          } else {
            cat("NOTE: No clinical data available in this custom version !\n")
            cat("\n")  
          }
        }
      }
    } else {
      cat("NOTE: No data available!\n")
      cat("\n")
    }
    
  }
}
```

# References
